---
output: github_document
editor_options:
  markdown:
    wrap: 72
---

# scholidonline

[![R-CMD-check](https://github.com/Thomas-Rauter/scholidonline/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Thomas-Rauter/scholidonline/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://img.shields.io/codecov/c/github/Thomas-Rauter/scholidonline?branch=main&logo=codecov)](https://app.codecov.io/gh/Thomas-Rauter/scholidonline)

`scholidonline` provides lightweight **online** utilities for working with
scholarly identifiers in R. It builds on `scholid` for identifier detection
and normalization, and adds minimal-dependency functions to query external
registries.

See the full documentation at the
[scholidonline website](https://thomas-rauter.github.io/scholidonline/).

## Installation

Install the released version from CRAN:

```{r install, eval = FALSE}
install.packages("scholidonline")
```

## Scope

The package focuses on online operations for common identifier systems used
in scholarly communication:

- DOI
- ORCID iD
- arXiv
- PubMed (PMID)
- PubMed Central (PMCID)

It provides registry-backed functionality such as:

- Existence checks
- Identifier conversion across systems
- Basic metadata retrieval
- Discovery of linked identifiers

## Interface

User-available functions:

| Function | Purpose |
|----------|----------|
| `id_exists(x, type = NULL)` | Check whether identifiers exist in their respective registries |
| `id_convert(x, to, from = NULL)` | Convert identifiers across systems (e.g., PMID → DOI) |
| `id_metadata(x, type = NULL)` | Retrieve basic structured metadata |
| `id_links(x, type = NULL)` | Discover linked identifiers |

All functions are vectorized and return predictable base R objects
(logical vectors or data.frames).

Identifier detection and normalization are delegated to `scholid`.

## Examples

```{r examples, eval = TRUE}
# check whether a DOI exists
scholidonline::id_exists(
  x    = "10.1000/182",
  type = "doi"
)

# infer identifier type automatically
scholidonline::id_exists(
  x = c("10.1000/182", "12345678")
)

# convert a PMID to a DOI
scholidonline::id_convert(
  x    = "12345678",
  from = "pmid",
  to   = "doi"
)

# retrieve metadata for a DOI
scholidonline::id_metadata(
  x    = "10.1000/182",
  type = "doi"
)

# discover linked identifiers for a PMCID
scholidonline::id_links(
  x    = "PMC1234567",
  type = "pmcid"
)
```

For more detailed usage patterns check out the **Get started** vignette.

## Relationship to scholid

`scholid` provides dependency-free utilities for detecting, normalizing,
classifying, and extracting scholarly identifiers.

`scholidonline` builds on that foundation and adds online registry queries,
while keeping dependencies minimal (`httr2`, `jsonlite`) and maintaining
CRAN-friendly behavior.

## License

MIT
