---
output: github_document
editor_options:
  markdown:
    wrap: 72
---

# scholidonline

[![R-CMD-check](https://github.com/Thomas-Rauter/scholidonline/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Thomas-Rauter/scholidonline/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://img.shields.io/codecov/c/github/Thomas-Rauter/scholidonline?branch=main&logo=codecov)](https://app.codecov.io/gh/Thomas-Rauter/scholidonline)

`scholidonline` provides lightweight **online** utilities for working with
scholarly identifiers in R. It builds on `scholid` for identifier detection
and normalization, and adds minimal-dependency functions to query external
registries.

See the full documentation at the
[scholidonline website](https://thomas-rauter.github.io/scholidonline/).

## Installation

Install the released version from CRAN:

```{r install, eval = FALSE}
install.packages("scholidonline")
```

## Scope

The package focuses on online operations for common identifier systems used
in scholarly communication:

- DOI
- ORCID iD
- arXiv
- PubMed (PMID)
- PubMed Central (PMCID)

It provides registry-backed functionality such as:

- Existence checks
- Identifier conversion across systems
- Basic metadata retrieval
- Discovery of linked identifiers

## Interface

User-available functions:

| Function | Purpose |
|----------|----------|
| `scholidonline_types` | Supported scholidonline identifier types |
| `scholidonline_capabilities` | Supported scholidonline capabilities |
| `id_exists(x, type = NULL)` | Check whether identifiers exist in their respective registries |
| `id_convert(x, to, from = NULL)` | Convert identifiers across systems (e.g., PMID → DOI) |
| `id_metadata(x, type = NULL)` | Retrieve basic structured metadata |
| `id_links(x, type = NULL)` | Discover linked identifiers |



## Examples

```{r examples, eval = TRUE}
# List supported scholidonline identifier types
scholidonline::scholidonline_types()

# List scholidonline capabilities
scholidonline::scholidonline_capabilities()

# Check if an ID exists online
scholidonline::id_exists(
  "10.1000/182",
  type = "doi"
  )

# Convert identifiers across systems
scholidonline::id_convert(
  "12345678",
  to = "doi",
  from = "pmid"
  )

# Retrieve scholarly metadata
out <- scholidonline::id_metadata(
  "10.1038/nature12373",
  type = "doi"
  )

# Show key fields
knitr::kable(out)

# Return linked scholarly identifiers
out <- scholidonline::id_links(
  "31452104",
  provider = "epmc"
  )

knitr::kable(out)
```



## Relationship to scholid

[`scholid`](https://thomas-rauter.github.io/scholid/) provides dependency-free
utilities for detecting, normalizing, classifying, and extracting scholarly
identifiers.

`scholidonline` builds on that foundation and adds online registry queries.



## License

MIT
